---
title: "Practical Machine Learning Course Assignment"
author: "Ashley Campisano"
date: "1/10/2020"
output: html_document
---

#Practical Machine Learning - Course Project

#Ashley Campisano

#Project Overview

#This project was done using activity data from Jawbone Up, Nike FuelBand, and Fitbit activity trackers. The goal of this project was to predict the manner in which the person from the traing set did the exercise (the "classe" variable). The data was from the accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

#Load R packages that may be needed for this project
```{r}
library(caret)
library(ggplot2)
library(randomForest)
library(gbm)
library(rpart)
library(rpart.plot)
library(corrplot)
library(rattle)
```
#Load the Training and Testing Data
```{r}
setwd("/Users/acampisano/Downloads")
train <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```
#Explore The Data
```{r}
names(train)
str(train)
summary(train)

summary(train$classe) #Classe is the result we want to predict using the other variables in the dataset
```
#Clean Up the Data

#Subset data for Cross Validation
```{r}
inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
myTrain <- train[inTrain, ]
myTest <- train[-inTrain, ]

dim(myTrain)
dim(myTest)
```
#Remove variables with more than 80% missing variables
```{r}
AllNA    <- sapply(myTrain, function(x) mean(is.na(x))) > 0.95
mytrain_SUB <- myTrain[, AllNA==FALSE]
mytest_SUB  <- myTest[, AllNA==FALSE]

dim(mytrain_SUB)
```
#Remove uneeded variables that are not predictors
```{r}
names(mytrain_SUB)
mytrain_SUB <- mytrain_SUB[,8:length(mytrain_SUB)]
```
#Remove near zero covariates
```{r}
nsv <- nearZeroVar(mytrain_SUB, saveMetrics = T)
mytrain_SUB <- mytrain_SUB[, !nsv$nzv]
mytest_SUB <- mytest_SUB[, !nsv$nzv]
```
#Prediction Modeling Methods:
#* Random Forest
#* Classification Trees
#* Generalized Boosted Model

#Random Forest

#Fit Model
```{r}
set.seed(1234)

modelFit <- randomForest(classe~., data = mytrain_SUB)
print(modelFit)
```
#Cross Validation
```{r}
predict_rf_cv <- predict(modelFit, myTest, type = "class")
confusionMatrix(myTest$classe, predict_rf_cv)
```
#Sample Error
```{r}
predict_rf_train <- predict(modelFit, myTrain, type = "class")
confusionMatrix(myTrain$classe, predict_rf_train)
```
#Classification Trees

#Fit Model
```{r}
set.seed(12345)
dtmodel <- rpart(classe ~ ., data = mytrain_SUB, method = "class")
fancyRpartPlot(dtmodel)
```
#Cross Validation
```{r}
predict_dt_cv <- predict(dtmodel, myTest, type = "class")
confusionMatrix(myTest$classe, predict_dt_cv)
```
#Sample Error
```{r}
predict_dt_train <- predict(dtmodel, myTrain, type = "class")
confusionMatrix(myTrain$classe, predict_dt_train)
```
#Generalized Boosted Model

#Fit Model
```{r}
set.seed(1234)

controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
Model_Fit_GBM  <- train(classe ~ ., data=mytrain_SUB, method = "gbm",trControl = controlGBM, verbose = FALSE)
Model_Fit_GBM
```
#Cross validation
```{r}
predict_gbm_cv <- predict(Model_Fit_GBM, myTest, type = "raw")
confusionMatrix(myTest$classe, predict_gbm_cv)
```
#Sample Error
```{r}
predict_gbm_train <- predict(Model_Fit_GBM, myTrain, type = "raw")
confusionMatrix(myTrain$classe, predict_gbm_train)
```
#Final Model - Decision

#Model Accuracy Comparison

#Prediction Model Accuracy
#* Random Forest: 0.9915 
#* Classification Trees: 0.749 
#* Generalized Boosted Model: 0.9592

#The Random Forest Model had the highest model accuracy with .9915 overall accuracy. Due to this, I chose to use my Random Forest MOdel for the final course assignment prediction.

#Final Model - Random Forest
```{r}
final_model <- predict(modelFit, test, type = "class")
print(final_model)

pml_write_files = function(x) {
  n = length(x)
  for (i in 1:n) {
    filename = paste0("problem_id_", i, ".txt")
    write.table(x[i], file=filename, quote=FALSE,row.names=FALSE, col.names=FALSE)
  }
}

pml_write_files(final_model)
```
